### 2016-07-19 mcmc.tex
renaming problems

\begin{problem}
Look up (or choose) definitions for the mean, variance, skewness,
and kurtosis of a distribution.
Also look up or compute the analytic values of these four statistics
for a top-hat (uniform) distribution.
Write a computer program that uses some standard package (such as
\project{numpy}\note{There isn't a full citation for this package but there is \cite{numpy}.})
to generate $K$ random numbers
$x$ from a uniform distribution in the interval $0<x<1$.
Now use those $K$ numbers to compute a the sampling estimate of the
mean, variance, skewness, and kurtosis (four estimates)
Make four plot of these four estimates as a function of $1/K$ or
perhaps $\log_2 K$, for $K=4^n$ for $n=1$ up to $n=10$ (that is,
$K=4$, $K=16$, and so on up to $K=1048576$).
Over-plot the analytic answers.
What can you conclude?
\end{problem}

\section{Likelihoods and priors}

\begin{problem}
Run your M-H \MCMC\ sampler from \problemname~\ref{prob:MH}, but now
with a density function that is precisely unity \emph{everywhere}
(that is, at any input value of $x$ it returns unity).  That is, an
improper function.  Run it for longer and longer and plot the chain
value $x$ as a function of timestep.  What happens?
\end{problem}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/p2}
\end{center}
\caption{Solution to \problemname~\ref{prob:MH}:
MCMC samples from a Gaussian (black) and the true distribution (blue).}
\label{fig:MH}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/p3}
\end{center}
\caption{Solution to \problemname~\ref{prob:MH2}:
MCMC samples from a uniform distribution (black) and the true distribution
(blue).}
\label{fig:MH2}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{problems/p4a}
\end{center}
\caption{Solution to \problemname~\ref{prob:twod}a:
MCMC samples from a two-dimensional Gaussian (scatter plot) and the
one-dimensional marginalized distributions (histograms).}
\label{fig:twod-a}
\end{figure}

@@

\begin{problem}
Why, in all the previous problems, did we give the proposal
distributions $q(x'\given x)$ zero mean?
What would be bad if we hadn't done that?
Re-do \problemname~\ref{prob:twod}a with a proposal $q(x'\given x)$ with a
non-zero mean and see what happens.
Bonus points: Modify the acceptance--rejection criterion to deal with
the messed-up $q(x'\given x)$ and show that everything works once
again.
\end{problem}

>>>>

\begin{problem}\label{prob:simple-stats}
Look up (or choose) definitions for the mean, variance, skewness,
and kurtosis of a distribution.
Also look up or compute the analytic values of these four statistics
for a top-hat (uniform) distribution.
Write a computer program that uses some standard package (such as
\project{numpy}\note{There isn't a full citation for this package but there is \cite{numpy}.})
to generate $K$ random numbers
$x$ from a uniform distribution in the interval $0<x<1$.
Now use those $K$ numbers to compute a the sampling estimate of the
mean, variance, skewness, and kurtosis (four estimates)
Make four plot of these four estimates as a function of $1/K$ or
perhaps $\log_2 K$, for $K=4^n$ for $n=1$ up to $n=10$ (that is,
$K=4$, $K=16$, and so on up to $K=1048576$).
Over-plot the analytic answers.
What can you conclude?
\end{problem}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/MH}
\end{center}
\caption{Solution to \problemname~\ref{prob:MH}:
MCMC samples from a Gaussian (black) and the true distribution (blue).}
\label{fig:MH}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{problems/MH2}
\end{center}
\caption{Solution to \problemname~\ref{prob:MH2}:
MCMC samples from a uniform distribution (black) and the true distribution
(blue).}
\label{fig:MH2}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{problems/twod_a}
\end{center}
\caption{Solution to \problemname~\ref{prob:twod}a:
MCMC samples from a two-dimensional Gaussian (scatter plot) and the
one-dimensional marginalized distributions (histograms).}
\label{fig:twod-a}
\end{figure}

@@

\begin{problem}\label{prob:prop-mean}
Why, in all the previous problems, did we give the proposal
distributions $q(x'\given x)$ zero mean?
What would be bad if we hadn't done that?
Re-do \problemname~\ref{prob:twod}a with a proposal $q(x'\given x)$ with a
non-zero mean and see what happens.
Bonus points: Modify the acceptance--rejection criterion to deal with
the messed-up $q(x'\given x)$ and show that everything works once
again.
\end{problem}

\section{Likelihoods and priors}\label{sec:likelihood}

\begin{problem}
Run your M-H \MCMC\ sampler from \problemname~\ref{prob:MH}, but now
with a density function that is precisely unity \emph{everywhere}
(that is, at any input value of $x$ it returns unity).  That is, an
improper function (as discussed in \sectionname~\ref{sec:likelihood}).
Run it for longer and longer and plot the chain value $x$ as a
function of timestep.  What happens?
\end{problem}
